### 2021-10-20 paper/output/document/paper41.tex
finished metaethical test

%
\begin{isabellebody}%
\setisabellecontext{paper{\isadigit{4}}{\isadigit{1}}}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Novel Formalization of the Categorical Imperative%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this section, I present a custom formalization of the categorical imperative, as inspired by 
the goals from the previous chapter.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Logical Background%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
The previous attempts to model the categorical imperative in Chapter 2 partially failed due to 
an inability to fully represent the complexity of a maxim. Specifically, they treated actions as a single, 
monolithic unit of evaluation, whereas most Kantians consider the unit of evaluation for the FUL to the more
complex notion of a maxim. In this section, I will present some logical background necessarily to fully 
capture the spirit of a maxim. I will begin by borrowing some machinery to handle ``subjects" who perform 
actions from Chapter 2.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{typedecl}\isamarkupfalse%
\ s\ %
\isamarkupcmt{s is the type for a ``subject," i.e. the subject of a sentence. In this interpretation, 
a subject is merely defined as ``that which can act." It does not include any other properties, such as 
rationality or dignity. As I will show, for the purposes of defining the universalizability test, this 
``thin" representation of a subject suffices.%
}\isanewline
\isanewline
\isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ os\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}s\ {\isasymRightarrow}\ t{\isacharparenright}{\isachardoublequoteclose}\ %
\isamarkupcmt{Recall that an open sentence maps a subject to a term to model the 
substitution operator.%
}\isanewline
\isanewline
\isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ maxim\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}t\ {\isacharasterisk}\ os\ {\isacharasterisk}\ t{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The central unit of evaluation for the universalizability test is a ``maxim," which Kant defines 
in a footnote in \emph{Groundworkd} as ``the subjective principle of willing," or the principle that 
the agent acts on $\cite[16]{groundwork}$. Modern Kantians differ in their interpretations of this definition. The naive view 
is that a maxim is an act, but Korsgaard adopts the more sophisticated view that a maxim is composed
of an act and the agent's purpose for acting \cite{actingforareason}. She also compares a maxim 
to Aristotle's logos, which includes these components and information about the circumstances and methods 
of the act. O'Neill concludes that Kant's examples imply that a maxim must also include circumstances \cite{actingonprinciple}, and 
Kitcher \cite{whatisamaxim} uses textual evidence from the Groundwork to argue for the inclusion of a maxim's purpose 
or motivation. In order to formalize the notion of a maxim, I must adopt a specific definition and 
defend my choice.

@@

While Kitcher's conception of a maxim captures Kant's idea of acting for duty's own sake, I will not implement it 
because it is not necessary for putting maxims through the FUL. Indeed, Kitcher acknowledges that 
O'Neill's formulation suffices for the universalizability test, but is not the general notion of a maxim.
In order to pass the maxim through the FUL, it suffices to know the circumstance, act, and goal. The FUL
derives the motive that Kitcher bundles into the maxim, so automating the FUL does not require 
including a motive. The ``input" to the FUL is the circumstance, act, goal tuple. My project takes 
this input and returns the motivation that the dutiful, moral agent would adopt. Additionally, doing
justice to the rich notion of motive requires modelling the operation of practical reason itself, 
which is outside the scope of this project. My work focuses on the universalizability test, but future work that 
models the process of practical reason may use my implementation of the FUL as a ``library." Combined 
with a logic of practical reason, an implementation of the FUL can move from evaluating a maxim to 
evaluating an agent's behavior, since that's when ``acting from duty" starts to matter.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ will\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}maxim\ {\isasymRightarrow}\ s{\isasymRightarrow}\ \ t{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}W\ {\isacharunderscore}\ {\isacharunderscore}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}will\ {\isasymequiv}\ {\isasymlambda}{\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s{\isachardot}\ {\isacharparenleft}c\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}a\ s{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{print{\isacharunderscore}theorems}\isamarkupfalse%
%
\begin{isamarkuptext}%
Korsgaard claims that ``to will an end, rather than just
wishing for it or wanting it, is to set yourself to be its cause" \cite{sources}. To will a maxim
is to set yourself to be the cause of its goal by taking the means 
specified in the maxim in the relevant circumstances. This coheres with 
Kitcher's and Korsgaard's understanding of a maxim as a principle or rule to live by.

@@

Worlds where the circumstances do not hold are not relevant for determining obligation. Recall that in 
\cite{BFP}'s definition of the obligation operator,  $O \{B|A\}$ is true at all worlds iff ob(B)(A), or 
if the obligation function maps A to obligatory in context B (where the context is a set of worlds). This 
definition implies that worlds outside of B have no bearing on the moral status of A in context B, which 
coheres with intuitions about contextual obligation. Thus, the dyadic obligation operator 
disqualifies worlds where the context does not hold, so the vacuous truth of the will statement in 
these worlds does not matter.

@@

@@

To get around this, I will exclude what I call ``badly formed maxims," which are those maxims such that the goal has already been 
achieved or the act has already been acted on. Under my formalization, such maxims are 
not well-formed. To understand why, I return to Korsgaard's and O'Neill's interpretations of a maxim as a practical
guide to action. A maxim is a practical principle that guides how we behave in everyday life. A 
principle of the form ``When you are eating breakfast, eat breakfast in order to eat breakfast," is not 
practically relevant. No agent would ever need to act on such a principle. It is not contradictory
or prohibited, but it is the wrong kind of question to be asking. It is not a 
well-formed maxim, so the categorical imperative does not apply to it.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ well{\isacharunderscore}formed{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}maxim{\isasymRightarrow}s{\isasymRightarrow}i{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}well{\isacharunderscore}formed\ {\isasymequiv}\ {\isasymlambda}{\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s\ w{\isachardot}\ {\isacharparenleft}{\isasymnot}\ {\isacharparenleft}\ {\isacharparenleft}c\ \isactrlbold {\isasymrightarrow}\ g{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isacharparenleft}{\isasymnot}\ {\isacharparenleft}\ {\isacharparenleft}c\ \isactrlbold {\isasymrightarrow}\ a\ s{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{This abbreviation formalizes the well-formedness of a maxim for a subject. The goal cannot be 
already achieved in the circumstances and the subject cannot have already performed the act.%
}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ FUL\ \isakeyword{where}\ {\isachardoublequoteopen}FUL\ {\isasymequiv}\ {\isasymforall}M{\isacharcolon}{\isacharcolon}maxim{\isachardot}\ {\isasymforall}s{\isacharcolon}{\isacharcolon}s{\isachardot}\ {\isacharparenleft}{\isasymforall}w{\isachardot}\ well{\isacharunderscore}formed\ M\ s\ w{\isacharparenright}\ {\isasymlongrightarrow}\ {\isacharparenleft}not{\isacharunderscore}universalizable\ M\ s\ {\isasymlongrightarrow}\ {\isasymTurnstile}\ prohibited\ M\ s\ {\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{Let's try the exact same formalization of the FUL as above, except that it only applies to 
maxims that are well-formed at every world.%
}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}FUL{\isachardoublequoteclose}\isanewline
\ \ \isacommand{nitpick}\isamarkupfalse%
{\isacharbrackleft}user{\isacharunderscore}axioms{\isacharcomma}\ falsify{\isacharequal}true{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\isanewline
%
\isamarkupcmt{The FUL does not hold in DDL, because nitpick is able to find a model for my system in which it is 
false. If the FUL were already a theorem of the system, adding it wouldn't make the system any more 
powerful, so this is the desired result.

@@

  Skolem constants:
    a = ($\lambda x. \_$)($s_1$ := ($\lambda x. \_$)($i_1$ := False))
    c = ($\lambda x. \_$)($i_1$ := True)
    g = ($\lambda x. \_$)($i_1$ := False)
    $\lambda w$. p = ($\lambda x. \_$)($i_1$ := $s_1$)
    s = $s_1$ $\color{black}$%
}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\ FUL{\isacharcolon}FUL\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{nitpick}\isamarkupfalse%
{\isacharbrackleft}user{\isacharunderscore}axioms{\isacharcomma}\ falsify{\isacharequal}false{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ simp\isanewline
%
\isamarkupcmt{Nitpick is able to find a model in which all axioms are satisfied, 
so this version of the FUL is consistent.

@@

The logic confirmed that certain kinds
of circumstance, act, goal tuples are too badly formed for the categorical imperative to logically 
apply to them. The realization of this subtle problem would have been incredibly difficult without 
computational tools. The syntax and typing of Isabelle/HOL forced me to bind the free-variable $M$
in the FUL in different ways and allowed me to quickly test many bindings. The discovery of this 
logical inconsistency then enabled a philosophical insight about which kinds of maxims make sense as 
practical principles. This is one way to do computational ethics: model a system in a logic, use 
computational tools to refine and debug the logic, and then use insights about the logic to derive 
insights about the ethical phenonema it is modelling. This procedure parallels the use of proofs in 
theoretical math to understand the mathematical objects they model.%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{paper{\isadigit{4}}{\isadigit{1}}}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Novel Formalization of the Categorical Imperative%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this section, I present a custom formalization of the categorical imperative, as inspired by 
the goals from the previous chapter.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Logical Background%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
The previous attempts to model the categorical imperative in Chapter 2 partially failed due to 
an inability to fully represent the complexity of a maxim. Specifically, they treated actions as a single, 
monolithic unit of evaluation, whereas most Kantians consider the unit of evaluation for the FUL to the more
complex notion of a maxim. In this section, I will present some logical background necessarily to fully 
capture the spirit of a maxim. I will begin by borrowing some machinery to handle ``subjects" who perform 
actions from Chapter 2.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{typedecl}\isamarkupfalse%
\ s\ %
\isamarkupcmt{s is the type for a ``subject," i.e. the subject of a sentence. In this interpretation, 
a subject is merely defined as ``that which can act." It does not include any other properties, such as 
rationality or dignity. As I will show, for the purposes of defining the universalizability test, this 
``thin" representation of a subject suffices.%
}\isanewline
\isanewline
\isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ os\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}s\ {\isasymRightarrow}\ t{\isacharparenright}{\isachardoublequoteclose}\ %
\isamarkupcmt{Recall that an open sentence maps a subject to a term to model the 
substitution operator.%
}\isanewline
\isanewline
\isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ maxim\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}t\ {\isacharasterisk}\ os\ {\isacharasterisk}\ t{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The central unit of evaluation for the universalizability test is a ``maxim," which Kant defines 
in a footnote in \emph{Groundwork} as ``the subjective principle of willing," or the principle that 
the agent acts on $\cite[16]{groundwork}$. Modern Kantians differ in their interpretations of this definition. The naive view 
is that a maxim is an act, but Korsgaard adopts the more sophisticated view that a maxim is composed
of an act and the agent's purpose for acting \cite{actingforareason}. She also compares a maxim 
to Aristotle's logos, which includes these components and information about the circumstances and methods 
of the act. O'Neill concludes that Kant's examples imply that a maxim must also include circumstances \cite{actingonprinciple}, and 
Kitcher \cite{whatisamaxim} uses textual evidence from the Groundwork to argue for the inclusion of a maxim's purpose 
or motivation. In order to formalize the notion of a maxim, I must adopt a specific definition and 
defend my choice.

@@

While Kitcher's conception of a maxim captures Kant's idea of acting for duty's own sake, I will not implement it 
because it is not necessary for putting maxims through the FUL. Indeed, Kitcher acknowledges that 
O'Neill's formulation suffices for the universalizability test, but is not the general notion of a maxim.
In order to pass the maxim through the FUL, it suffices to know the circumstance, act, and goal. The FUL
derives the motive that Kitcher bundles into the maxim, so automating the FUL does not require 
including a motive. The ``input" to the FUL is the circumstance, act, goal tuple. My project takes 
this input and returns the motivation that the dutiful, moral agent would adopt. Additionally, doing
justice to the rich notion of motive requires modelling the operation of practical reason itself, 
which is outside the scope of this project. My work focuses on the universalizability test, but future work that 
models the process of practical reason may use my implementation of the FUL as a ``library." Combined 
with a logic of practical reason, an implementation of the FUL can move from evaluating a maxim to 
evaluating an agent's behavior, since that's when ``acting from duty" starts to matter.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ will\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}maxim\ {\isasymRightarrow}\ s{\isasymRightarrow}\ \ t{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}W\ {\isacharunderscore}\ {\isacharunderscore}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}will\ {\isasymequiv}\ {\isasymlambda}{\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s{\isachardot}\ {\isacharparenleft}c\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}a\ s{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{print{\isacharunderscore}theorems}\isamarkupfalse%
%
\begin{isamarkuptext}%
Korsgaard claims that ``to will an end, rather than just
wishing for it or wanting it, is to set yourself to be its cause" \cite[38]{sources}. To will a maxim
is to set yourself to be the cause of its goal by taking the means 
specified in the maxim in the relevant circumstances. This coheres with 
Kitcher's and Korsgaard's understanding of a maxim as a principle or rule to live by.

@@

Worlds where the circumstances do not hold are not relevant for determining obligation. Recall that in 
Benzmueller et. al's definition of the obligation operator,  $O \{B|A\}$ is true at all worlds iff ob(B)(A), or 
if the obligation function maps A to obligatory in context B (where the context is a set of worlds) \cite{BFP}. This 
definition implies that worlds outside of B have no bearing on the moral status of A in context B, which 
coheres with intuitions about contextual obligation. Thus, the dyadic obligation operator 
disqualifies worlds where the context does not hold, so the vacuous truth of the will statement in 
these worlds does not matter.

@@

As before, the concepts of prohibition and permissibility will be helpful here. The unit of 
evaluation for my formalization of the FUL is the act of willing a maxim, which entails performing 
the maxim's act in the relevant circumstances. Therefore, I will say that, just as the act of willing a
 maxim can be obligatory for a subject, it can be prohibited or permissible for a subject.\footnote{In 
the rest of this section, for convenience, I will use the phrase ``subject s willing maxim M is obligatory" 
interchangeably with ``maxim M is obligatory for subject s." I will use ``maxim M is obligatory" to 
refer to M being obligatory for any arbitrary subject, which I will show to be equivalent to M being 
obligatory for a specific subject.}%
\ permissible{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}maxim{\isasymRightarrow}s{\isasymRightarrow}t{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{where}\ {\isachardoublequoteopen}permissible\ {\isasymequiv}\ {\isasymlambda}M\ s{\isachardot}\ \isactrlbold {\isasymnot}\ {\isacharparenleft}prohibited\ M\ s{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isamarkupcmt{I will say that a maxim is permissible for a subject if it is not prohibited for that subject to 
will that maxim.%
One problem with prior formalization of the categorical imperative was that they didn't
prohibit contradictory obligations, partially because DDL itself allows contradictory obligations. 
Kant subscribes to the general, popular view that morality is supposed to guide action, so ought implies 
can\footnote{Kohl points out that this principle is referred to as 
Kant's dictum or Kant's law in the literature. \cite[footnote 1]{kohl}}. Kohl reconstructs his argument for the principle as 
follows: if the will cannot comply with the moral law, then the moral law has no prescriptive authority 
for the will \cite[703-4]{kohl}. This defeats the purpose of Kant's theory—to develop an unconditional, categorical imperative 
for rational agents. Ought implies can requires that obligations never contradict, because an agent 
can't perform contradictory actions. Therefore, any ethical theory that respects ought implies can, 
and Kantian ethics in particular, must not result in conflicting obligations.

Kant only briefly discusses contradictory obligations in \emph{Metaphysics of Morals}, where he argues that 
conflicting moral obligations are impossible under his theory \cite[V224]{metaphysicsintro}. Particularly, the categorical imperative generates 
``strict negative laws of omission", which cannot conflict by definition \cite[45]{timmerman}.\footnote{The 
kinds of obligations generated by the FUL are called ``perfect duties" which arise from ``contradictions 
in conception," or maxims that we cannot even concieve of universalizing. These duties are always negative 
and thus never conflict. Kant also presents ``imperfect duties," generated from ``contradictions in will,"
or maxims that we can concieve of universalizing but would never want to. These duties tend to be broader, 
such as ``improve oneself" or "help others," and are secondary to perfect duties. My project only analyzes 
perfect duties, as these are stronger than imperfect duties.}

When analyzing the naive formalization and Kroy's formalization, I learned that DDL and the prior 
formalizations allow contradictory obligations. This is a major weakness of these systems, and my 
formalization should fix this. To do so, I will add as an axiom the idea that obligations cannot 
contradict each other or their internal circumstances. Formally, conflicting obligations are defined below.%
\ non{\isacharunderscore}contradictory\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}non{\isacharunderscore}contradictory\ A\ B\ c\ w\ {\isasymequiv}\ {\isacharparenleft}{\isacharparenleft}O{\isacharbraceleft}A{\isacharbar}c{\isacharbraceright}\ \isactrlbold {\isasymand}\ O{\isacharbraceleft}B{\isacharbar}c{\isacharbraceright}{\isacharparenright}\ w{\isacharparenright}\ {\isasymlongrightarrow}\ {\isasymnot}{\isacharparenleft}{\isacharparenleft}A\ \isactrlbold {\isasymand}\ {\isacharparenleft}B\ \isactrlbold {\isasymand}\ c{\isacharparenright}{\isacharparenright}\ w\ {\isasymlongrightarrow}\ False{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isamarkupcmt{Terms A and B are non contradictory in circumstances c if, when A and B are obligated in circumstances 
c, the conjunction of A, B, and c, does not imply False.%
\isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\ no{\isacharunderscore}contradictions{\isacharcolon}{\isachardoublequoteopen}{\isasymforall}A{\isacharcolon}{\isacharcolon}t{\isachardot}\ {\isasymforall}B{\isacharcolon}{\isacharcolon}t{\isachardot}\ {\isasymforall}c{\isacharcolon}{\isacharcolon}t{\isachardot}\ {\isasymforall}w{\isacharcolon}{\isacharcolon}i{\isachardot}\ non{\isacharunderscore}contradictory\ A\ B\ c\ w{\isachardoublequoteclose}\isanewline
\isamarkupcmt{This axiom formalizes the idea that, for any terms A, B, and circumstances c, A and B must be 
non-contradictory in circumstances c at all worlds. Intuitively, this axiom requires that obligations 
do not conflict.%
\isadelimdocument
\endisadelimdocument
\isatagdocument
\isamarkupsubsection{Formalizing the FUL%
}
\isamarkuptrue%
\endisatagdocument
{\isafolddocument}%
\isadelimdocument
\endisadelimdocument

@@

To get around this, I will exclude what I call ``badly formed maxims," which are those maxims such that the goal has already been 
achieved or the act has already been acted on. Under my formalization, such maxims are 
not well-formed. To understand why, I return to Korsgaard's and O'Neill's interpretations of a maxim as a practical
guide to action. A maxim is a practical principle that guides how we behave in everyday life. A 
principle of the form ``When you are eating breakfast, eat breakfast in order to eat breakfast," is not 
practically relevant. No agent would ever need to act on such a principle. It is not contradictory
or prohibited, but it is the wrong kind of question to be asking. It is not a 
well-formed maxim, so the categorical imperative does not apply to it. (more explanation in philosophical 
writing collection)%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ well{\isacharunderscore}formed{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}maxim{\isasymRightarrow}s{\isasymRightarrow}i{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ \isanewline
{\isachardoublequoteopen}well{\isacharunderscore}formed\ {\isasymequiv}\ {\isasymlambda}{\isacharparenleft}c{\isacharcomma}\ a{\isacharcomma}\ g{\isacharparenright}\ s\ w{\isachardot}\ {\isacharparenleft}{\isasymnot}\ {\isacharparenleft}\ {\isacharparenleft}c\ \isactrlbold {\isasymrightarrow}\ g{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isacharparenleft}{\isasymnot}\ {\isacharparenleft}\ {\isacharparenleft}c\ \isactrlbold {\isasymrightarrow}\ a\ s{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{This abbreviation formalizes the well-formedness of a maxim for a subject. The goal cannot be 
already achieved in the circumstances and the subject cannot have already performed the act.%
}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ FUL\ \isakeyword{where}\ {\isachardoublequoteopen}FUL\ {\isasymequiv}\ {\isasymforall}M{\isacharcolon}{\isacharcolon}maxim{\isachardot}\ {\isasymforall}s{\isacharcolon}{\isacharcolon}s{\isachardot}\ {\isacharparenleft}{\isasymforall}w{\isachardot}\ well{\isacharunderscore}formed\ M\ s\ w{\isacharparenright}\ {\isasymlongrightarrow}\ {\isacharparenleft}not{\isacharunderscore}universalizable\ M\ s\ {\isasymlongrightarrow}\ {\isasymTurnstile}\ prohibited\ M\ s\ {\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{Let's try the exact same formalization of the FUL as above, except that it only applies to 
maxims that are well-formed at every world.%
}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}FUL{\isachardoublequoteclose}\isanewline
\ \ \isacommand{nitpick}\isamarkupfalse%
{\isacharbrackleft}user{\isacharunderscore}axioms{\isacharcomma}\ falsify{\isacharequal}true{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\isanewline
%
\isamarkupcmt{The FUL does not hold in DDL, because nitpick is able to find a model for my system in which it is 
false. If the FUL were already a theorem of the system, adding it wouldn't make the system any more 
powerful, so this is the desired result.

@@

  Skolem constants:
    M = (($\lambda x. \_$)($i_1$ := True), ($\lambda x. \_$)($s_1$ := ($\lambda x. \_$)($i_1$ := False)), ($\lambda x. \_$)
         ($i_1$ := False))
    $\lambda$ w. p = ($\lambda x. \_$)($i_1$ := $s_1$)
    s = $s_1$ $\color{black}$%
}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\ FUL{\isacharcolon}FUL\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{nitpick}\isamarkupfalse%
{\isacharbrackleft}user{\isacharunderscore}axioms{\isacharcomma}\ falsify{\isacharequal}false{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ simp\isanewline
%
\isamarkupcmt{Nitpick is able to find a model in which all axioms are satisfied, 
so this version of the FUL is consistent.

@@

The logic confirmed that certain kinds
of circumstance, act, goal tuples are too badly formed for the categorical imperative to logically 
apply to them. The realization of this subtle problem would have been incredibly difficult without 
computational tools. The syntax and typing of Isabelle/HOL forced me to bind the free-variable $M$
in the FUL in different ways and allowed me to quickly test many bindings. The discovery of this 
logical inconsistency then enabled a philosophical insight about which kinds of maxims make sense as 
practical principles. This is one way to do computational ethics: model a system in a logic, use 
computational tools to refine and debug the logic, and then use insights about the logic to derive 
insights about the ethical phenonema it is modelling. This procedure parallels the use of proofs in 
theoretical math to understand the mathematical objects they model.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
One potential problem with my formalization is that it does not use the modal nature of the system. 
All of the properties that the FUL investigates hold at all worlds, in effect removing the modal nature 
of the system. This approach simplifies logical and therefore computational complexity, improving 
performance. On the other hand, it doesn't use the full expressivity of DDL. If I run into problems 
later on, one option is to tweak the FUL to use this expressivity.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Application Tests%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As with the naive formalization and Kroy's formalization, I will apply my testing framework to 
my custom formalization of the FUL. I will begin with some basic application tests. In these tests, 
I specify particular maxims as constants with no properties and gradually add properties to understand 
how the system handles different kinds of maxims.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Metaethical Tests%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Recall that metaethical tests test formal properties of the system that apply to any maxim, not 
just those specified in the application tests. In this section I adapt the metaethical tests developed 
in previous sections to my formalization of the categorical imperative. I preserved the philosophical 
goal of each test but modified them to test the stronger, richer notion of a maxim.

The first set of tests consider how obligation generalizes, first across worlds and then across
people. As expected, the tests below show that both wrongness (prohibition) and rightness (obligation)
generalize across both worlds and people. In other words, if something is obligated at some world, it 
is obligated at every world and if something is obligated for some person, then it is obligated for 
every person.

Generalization across people coheres with Kantian ethics—maxims are not person-specific.
Indeed, Velleman argues that, because reason is accessible to everyone identically, obligations apply 
to all people \cite[25]{velleman}. When Kant describes the categorical imperative as the objective 
principle of the will, he is referring to the fact that, as opposed to a subjective principle, the categorical
imperative applies to all rational agents equally $\cite[16]{groundwork}$.

Generalization across worlds is a consequence of the fact that my interpretation does not make use of the 
modal nature of DDL. In particular, I do not use any property of the world when prescribing obligations 
at that world.%
\ wrong{\isacharunderscore}if{\isacharunderscore}wrong{\isacharunderscore}for{\isacharunderscore}someone{\isacharcolon}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}w{\isachardot}\ {\isasymforall}c{\isacharcolon}{\isacharcolon}t{\isachardot}\ {\isasymforall}g{\isacharcolon}{\isacharcolon}t{\isachardot}\ {\isasymexists}s{\isacharcolon}{\isacharcolon}s{\isachardot}\ O{\isacharbraceleft}\isactrlbold {\isasymnot}\ {\isacharparenleft}W\ {\isacharparenleft}c{\isacharcomma}\ M{\isacharcomma}\ g{\isacharparenright}\ s{\isacharparenright}\ {\isacharbar}\ c{\isacharbraceright}\ w\ {\isasymlongrightarrow}\ {\isacharparenleft}{\isasymforall}p{\isachardot}\ \ O{\isacharbraceleft}\isactrlbold {\isasymnot}\ {\isacharparenleft}W\ {\isacharparenleft}c{\isacharcomma}\ M{\isacharcomma}\ g{\isacharparenright}\ p{\isacharparenright}\ {\isacharbar}\ c{\isacharbraceright}\ w{\isacharparenright}\ {\isachardoublequoteclose}\isanewline
\ \ %
\isacommand{by}\isamarkupfalse%
\ blast%
\ right{\isacharunderscore}if{\isacharunderscore}right{\isacharunderscore}for{\isacharunderscore}someone{\isacharcolon}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}w{\isachardot}\ {\isasymforall}c{\isacharcolon}{\isacharcolon}t{\isachardot}\ {\isasymforall}g{\isacharcolon}{\isacharcolon}t{\isachardot}\ {\isasymexists}s{\isacharcolon}{\isacharcolon}s{\isachardot}\ O{\isacharbraceleft}W\ {\isacharparenleft}c{\isacharcomma}\ M{\isacharcomma}\ g{\isacharparenright}\ s\ {\isacharbar}\ c{\isacharbraceright}\ w\ {\isasymlongrightarrow}\ {\isacharparenleft}{\isasymforall}p{\isachardot}\ \ O{\isacharbraceleft}W\ {\isacharparenleft}c{\isacharcomma}\ M{\isacharcomma}\ g{\isacharparenright}\ p\ {\isacharbar}\ c{\isacharbraceright}\ w{\isacharparenright}\ {\isachardoublequoteclose}\isanewline
\ \ %
\isacommand{by}\isamarkupfalse%
\ blast%
\ wrong{\isacharunderscore}if{\isacharunderscore}wrong{\isacharunderscore}somewhere{\isacharcolon}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}c\ g{\isachardot}\ {\isasymexists}w{\isadigit{1}}{\isachardot}\ O{\isacharbraceleft}\isactrlbold {\isasymnot}\ {\isacharparenleft}W\ {\isacharparenleft}c{\isacharcomma}\ M{\isacharcomma}\ g{\isacharparenright}\ s{\isacharparenright}\ {\isacharbar}\ c{\isacharbraceright}\ w{\isadigit{1}}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\isasymforall}w{\isadigit{2}}{\isachardot}\ \ O{\isacharbraceleft}\isactrlbold {\isasymnot}\ {\isacharparenleft}W\ {\isacharparenleft}c{\isacharcomma}\ M{\isacharcomma}\ g{\isacharparenright}\ s{\isacharparenright}\ {\isacharbar}\ c{\isacharbraceright}\ w{\isadigit{2}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ %
\isacommand{by}\isamarkupfalse%
\ blast%
\ right{\isacharunderscore}if{\isacharunderscore}right{\isacharunderscore}somewhere{\isacharcolon}\isanewline
